---
title: "Corridor TSP — Objective Comparison"
subtitle: "Kelvin Grove: Z₁ Delay · Z₂ Progression · Z₃ Lateness"
author: "Batch Runner"
format: 
  revealjs:
    theme: dark
    slide-number: true
    chalkboard: false
    width: 1400
    height: 900
    margin: 0.05
    center: false
execute:
  echo: false
  warning: false
---

```{r setup, include=FALSE}
library(tidyverse); library(scales); library(gt); library(patchwork)
theme_set(theme_minimal(base_size=13) +
  theme(plot.background=element_rect(fill="#1a1a2e",color=NA),
        panel.background=element_rect(fill="#16213e"),
        text=element_text(color="#e0e0e0"), axis.text=element_text(color="#a0a0a0",size=10),
        panel.grid=element_line(color="#2a2a4a"), legend.background=element_rect(fill="#16213e"),
        legend.text=element_text(color="#e0e0e0"),
        plot.title=element_text(size=14,face="bold"), plot.subtitle=element_text(size=10,color="#8888aa")))

batch <- read_csv("batch_results.csv", show_col_types=FALSE)
df <- batch %>% rename(experiment=run_experiment,
  pass_delay_hrs=stats_TotalPassDelay_hrs, main_delay_hrs=stats_MainPassDelay_hrs,
  side_delay_hrs=stats_SidePassDelay_hrs, bus_delay_s=stats_AvgBusPassDelay_s,
  car_delay_s=stats_AvgCarPassDelay_s, speed_kmh=stats_Net_AvgSpeed_kmh,
  detections=stats_TSP_Detections, extensions=stats_TSP_Extensions,
  insertions=stats_TSP_Insertions, alpha=wobj_alpha, beta=wobj_beta,
  w_main=wobj_w_main, w_side=wobj_w_side, rho_bus=wobj_rho_bus, rho_car=wobj_rho_car) %>%
  mutate(grant_rate=(extensions+insertions)/pmax(detections,1))

no_tsp <- df %>% filter(experiment=="NO_TSP")
base_delay <- no_tsp$pass_delay_hrs[1]; base_bus <- no_tsp$bus_delay_s[1]; base_car <- no_tsp$car_delay_s[1]

strategies <- df %>% filter(experiment %in% c("NO_TSP","DCTSP_MARL","DCTSP_ZIG",
         "DCTSP_BARGAIN_SPM","DCTSP_MP_ECTM","DCTSP_BXT")) %>%
  mutate(strategy_label=case_when(experiment=="NO_TSP"~"NoPriority",
    experiment=="DCTSP_MARL"~"CPD-QL", experiment=="DCTSP_ZIG"~"WaveGate",
    experiment=="DCTSP_BARGAIN_SPM"~"NashGate",
    experiment=="DCTSP_MP_ECTM"~"CellSearch", experiment=="DCTSP_BXT"~"CellQLearn"),
    vs_baseline=pass_delay_hrs-base_delay)

# Check if BUS_PRIO data exists
busprio_real <- df %>% filter(str_detect(experiment, "BUSPRIO_SWEEP"))
has_busprio <- nrow(busprio_real) > 0

bus_trips_path <- "results/NO_TSP_seed300_11129236_11129237_11129240/bus_trips.csv"
if(file.exists(bus_trips_path)) {
  bus_tt <- read_csv(bus_trips_path,show_col_types=FALSE)$TravelTime_s
  bus_tt <- bus_tt[!is.na(bus_tt)&bus_tt>0]
  sched_tt <- as.numeric(quantile(bus_tt,0.10)); lateness <- bus_tt-sched_tt
}
pct_30 <- if(exists("lateness")) round(mean(lateness>30)*100,0) else "?"
```

# What We're Comparing

::: {.notes}
**Presenter notes:**

We're testing 6 strategies on the Kelvin Grove corridor (12 intersections, Brisbane).
All strategies solve the same problem: should we grant TSP to an approaching bus,
and if so, which action (extend green / insert a phase / do nothing / ...)?

The difference is HOW they decide — and whether a **gate** checks the decision
before it's applied:

- **NoPriority**: never grants TSP — baseline, no gate.
- **CPD-QL**: Q-learning agent picks an action from learned experience — no
  gate, so a bad action is applied even when it hurts cross-traffic.
- **WaveGate**: simulates the corridor with an LWR shockwave model and grants
  only if the predicted bus-side pax·s saving exceeds the car-side pax·s cost.
- **CellSearch**: simulates ALL 10 candidate actions on a Cell Transmission
  Model (CTM — see below), picks the lowest-cost one, and grants it only if
  it beats doing nothing by enough.
- **CellQLearn**: same CTM simulation as CellSearch, but a Q-learning policy
  (not exhaustive search) picks which action to evaluate/apply.
- **NashGate**: Nash bargaining — picks the action that maximises
  (bus benefit) × (car benefit); an action that helps one side a lot but hurts
  the other too much scores low and is rejected.
- **META-TSP**: CellSearch's CTM search, PLUS (a) an adaptive bus weight that
  grows the more behind-schedule the bus already is, and (b) a hard
  "lateness takeover" — if a bus is critically late, priority is granted
  unconditionally, bypassing the cost-benefit check entirely.

**What is "the gate"?** Every strategy except CPD-QL has a gate: a pass/fail
check, run fresh at every bus detection, that decides whether to actually
grant priority. The methods differ in HOW they estimate "is this action
good?" (a shockwave model for WaveGate, a CTM grid search for
CellSearch/CellQLearn/META-TSP, a bargaining objective for NashGate) and WHAT
threshold they apply.

**The CTM gate (CellSearch, CellQLearn, META-TSP):** CTM = Cell Transmission
Model — a fast "what-if" simulator that splits the corridor into ~100m cells
and propagates traffic forward a short horizon for each candidate action (do
nothing / extend green ×1-3 / insert a green phase ×1-3 / ...). Each action
gets a cost = weighted bus delay + cross-traffic delay. The gate picks the
lowest-cost action $a^*$ and grants it only if
$Z(\text{NO\_ACTION}) - Z(a^*) \ge$ threshold AND the bus is already delayed
by enough. META-TSP adds: the bus-delay weight grows the later the bus
already is (adaptive weight), and if the bus is critically late
($|\sigma| \ge \tau$), the gate is skipped entirely and priority is granted no
matter what the CTM says (lateness takeover).

We measure 3 objectives:
- **Z₁**: total passenger delay (bus + car, in pax·s)
- **Z₂**: green-wave progression quality
- **Z₃**: bus lateness (headway deviation)

Key metric: **total passenger delay** aggregates across ALL ~94,700 passengers.
Vehicle-hours counts vehicles. Passenger delay is ~4× larger because buses carry 40 people.
:::

| Strategy | How it decides | Gate |
|----------|---------------|------|
| **NoPriority** | Never grants TSP | — |
| **CPD-QL** | Q-learning (54 states, ε-greedy) | No |
| **WaveGate** | LWR shockwave: grant if bus pax·s saved > car pax·s cost | Yes |
| **CellSearch** | CTM grid search: simulate 10 actions, grant the best if it beats doing nothing | Yes |
| **CellQLearn** | CTM grid search, but Q-learning picks the action | Yes |
| **NashGate** | Nash bargaining: maximise bus-benefit × car-benefit | Yes |
| **META-TSP** | CTM grid search + adaptive bus weight + lateness takeover | Yes (3-part) |

---

# Strategy Performance

```{r}
strategies %>% arrange(pass_delay_hrs) %>%
  mutate(Total=pass_delay_hrs, Delta=pass_delay_hrs-base_delay, Main=main_delay_hrs, Side=side_delay_hrs,
         Bus=bus_delay_s, Car=car_delay_s, Label=strategy_label) %>%
  select(Label, Total, Delta, Main, Side, Bus, Car, grant_rate) %>%
  gt() %>% tab_header(title="Total Passenger Delay", subtitle=paste0("NoPriority baseline: ",round(base_delay,1)," pax·hrs  |  all 6 strategies, real data")) %>%
  fmt_number(columns=c(Total,Delta,Main,Side,Bus,Car),decimals=1) %>% fmt_percent(columns=grant_rate,decimals=0) %>%
  cols_label(Label="", Total="Total (h)", Delta="Δ (h)", Main="Main (h)", Side="Side (h)", Bus="Bus (s)", Car="Car (s)", grant_rate="Grant") %>%
  tab_options(table.font.size="85%", data_row.padding="2px") %>%
  data_color(columns=Delta, fn=function(x) scales::col_numeric("RdYlGn",domain=range(x,na.rm=TRUE),reverse=TRUE)(x))
```

::: {.notes}
**How to read this table:**

- **Total**: passenger delay in pax·hours. Sum of every passenger's extra waiting time.
  NoPriority = 1261h means 94,730 passengers each delayed ~48s on average.
- **Δ**: change from NoPriority. Green = better (negative), red = worse (positive).
- **Main/Side**: the Total split by corridor section — Main = the coordinated
  TSP corridor, Side = the cross/side streets. Main + Side = Total.
- **Bus/Car**: per-person delay. Bus at 72.6s means each bus passenger waits 73 extra seconds.
- **Grant**: % of bus detections where TSP fired.

**Key story:**
- CPD-QL makes things WORSE (+509 pax·hrs). It has no guard, so it grants TSP 50% of the time,
  disrupting the green wave. Car delay jumps from 36→71 s/pax.
- WaveGate and NashGate both IMPROVE on baseline because their guards block harmful actions.
- NashGate has lowest grant rate (25%) — it's the most selective.
- **Main/Side reveals WHERE the win comes from**: relative to NoPriority's
  394.9h Main / 866.1h Side split, every strategy INCREASES Main delay
  (+38 to +589h) but most strategies DECREASE Side delay (by up to -328h
  for NashGate). The three Pareto-improving strategies (WaveGate, CellSearch,
  NashGate) all trade a modest Main-corridor cost for a much larger Side
  saving — the net win comes almost entirely from clearing side-street
  queues. CellQLearn is the exception: it gets WORSE on both Main (+149h)
  and Side (+84h).
:::

---

# Total Vehicle-Hours, Volume & Average Travel Time

```{r}
base_row <- strategies %>% filter(experiment=="NO_TSP")
base_total_tt <- with(base_row, stats_Net_TotalTT_h_Bus+stats_Net_TotalTT_h_Car+stats_Net_TotalTT_h_Truck)
base_vehicles <- with(base_row, stats_N_DistinctCars+stats_N_DistinctBuses+stats_N_DistinctTrucks)
base_avgtt <- base_total_tt*3600/base_vehicles

strategies %>% arrange(pass_delay_hrs) %>%
  mutate(BusTT=stats_Net_TotalTT_h_Bus, CarTT=stats_Net_TotalTT_h_Car, TruckTT=stats_Net_TotalTT_h_Truck,
         TotalTT=BusTT+CarTT+TruckTT,
         TotalVeh=stats_N_DistinctCars+stats_N_DistinctBuses+stats_N_DistinctTrucks,
         AvgTT_all=TotalTT*3600/TotalVeh, DeltaAvgTT=AvgTT_all-base_avgtt,
         Car_count=stats_N_DistinctCars, AvgCarTT=round(CarTT/Car_count*3600,0),
         AvgBusTT=stats_AvgBusTT_s) %>%
  select(strategy_label, BusTT, CarTT, TruckTT, TotalTT, TotalVeh, AvgTT_all, DeltaAvgTT, AvgBusTT, AvgCarTT, speed_kmh) %>%
  gt() %>% tab_header(title="Vehicle-Hours, Vehicle Volume & Average Travel Time (1h simulation)",
    subtitle=paste0("NoPriority: ",round(base_total_tt,1)," veh·hrs, ",base_vehicles," vehicles, ",round(base_avgtt,1),"s/vehicle")) %>%
  fmt_number(columns=c(BusTT,CarTT,TruckTT,TotalTT,AvgTT_all,DeltaAvgTT,speed_kmh,AvgBusTT,AvgCarTT),decimals=1) %>%
  fmt_number(columns=TotalVeh,decimals=0) %>%
  cols_label(strategy_label="", BusTT="Bus (h)", CarTT="Car (h)", TruckTT="Truck (h)", TotalTT="Total (h)",
             TotalVeh="Total Vehicles", AvgTT_all="Avg TT all (s)", DeltaAvgTT="Δ Avg TT (s)",
             AvgBusTT="Avg Bus TT (s)", AvgCarTT="Avg Car TT (s)", speed_kmh="km/h") %>%
  tab_options(table.font.size="75%", data_row.padding="2px") %>%
  data_color(columns=DeltaAvgTT, fn=function(x) scales::col_numeric("RdYlGn",domain=range(x,na.rm=TRUE),reverse=TRUE)(x))
```

::: {.notes}
**How to read:**
- **Vehicle-hours** (Bus/Car/Truck/Total) sum each vehicle's journey time, not weighted by
  occupancy — a bus and a car each contribute 1 vehicle-hour per hour in the network, even
  though the bus carries ~40 passengers. Total (≈316h for NoPriority) is therefore ~4×
  smaller than Total pax·hrs on the Strategy Performance slide (1261h) — that ratio IS the
  occupancy weighting (ρ_bus=40).
- **Total Vehicles** = distinct cars + buses + trucks completing a trip in the hour. Varies
  8134–8620 across strategies — NOT because demand differs (same seed for all 60 executed runs), but
  because congestion changes how many vehicles clear the network in time.
- **Avg TT all** = Total veh·hrs × 3600 / Total Vehicles — average journey time across EVERY
  vehicle, regardless of class.
- **Δ Avg TT** = increase over NoPriority's Avg TT all (131.9s) — the average extra delay
  per vehicle each strategy imposes.

**Avg Bus TT / Avg Car TT**: per-class average journey times (same as previous slides).
NoPriority bus=44.9s, car=126s. CPD-QL bus=63.2s, car=239s (nearly doubled!).

**Key story:**
- CPD-QL: Δ Avg TT = +117.5s — every vehicle takes ~2 minutes longer on average. Total +255.7 veh·hrs.
- CellSearch has the SMALLEST Δ Avg TT of any TSP strategy (+21.9s) — smaller than WaveGate
  (+25.5s) and NashGate (+28.9s). Its benefit is spread evenly across vehicles.
- NashGate adds 62.8 veh·hrs total but has the best car protection (see next slides).

**Why does Δ Avg TT go UP here while total passenger delay (previous slide)
goes DOWN — isn't that a contradiction?** No — they're different weightings
of the same green-time reallocation. TSP takes green time from cars and gives
it to buses. The previous slide's "Total (pax·hrs)" weights each bus-second
~27× a car-second (ρ_bus=40 vs ρ_car=1.5), so the (few) bus-seconds saved
outweigh the (many) car-seconds lost, and Total/Z₁ falls. **Δ Avg TT here is
UNWEIGHTED**: it averages the change across all ~8,300–8,600 vehicles, where
a few dozen buses each saving tens of seconds are outnumbered roughly 200:1
by cars each losing a few seconds — so the aggregate car-side cost (invisible
to Z₁) dominates, and Δ Avg TT rises for *every* TSP configuration tested.
In short: the previous slide answers "did the average **passenger** get a
better deal?" (yes); this slide answers "did the average **vehicle**'s trip
get faster?" (no, slightly slower). Both are correct simultaneously.
:::

---

# Bus vs Car Delay

```{r fig.height=6, fig.width=10}
strategies %>% filter(experiment!="NO_TSP") %>%
  mutate(lbl_vjust=case_when(strategy_label=="WaveGate" ~ -1.2,
                              strategy_label=="CellSearch" ~ 1.6,
                              strategy_label=="NashGate" ~ 1.4,
                              TRUE ~ -1.2),
         lbl_hjust=case_when(strategy_label=="WaveGate" ~ -0.1,
                              strategy_label=="NashGate" ~ -0.1,
                              TRUE ~ 0.5)) %>%
  ggplot(aes(bus_delay_s,car_delay_s,label=strategy_label)) +
  annotate("rect",xmin=-Inf,xmax=base_bus,ymin=-Inf,ymax=base_car,fill="#27ae60",alpha=0.08) +
  geom_vline(xintercept=base_bus,linetype="dashed",color="#3498db",alpha=0.5) +
  geom_hline(yintercept=base_car,linetype="dashed",color="#e67e22",alpha=0.5) +
  annotate("text",x=base_bus-4,y=base_car-2,label="Both better",color="#27ae60",size=3.5,hjust=1) +
  geom_point(aes(size=grant_rate,color=pass_delay_hrs),alpha=0.9) +
  geom_text(aes(vjust=lbl_vjust,hjust=lbl_hjust),color="#e0e0e0",size=3.8) +
  scale_color_viridis_c(name="Total (h)") + scale_size_continuous(name="Grant",labels=percent_format()) +
  labs(x="Bus Delay (s/pax)",y="Car Delay (s/pax)") + theme(legend.position="right")
```

::: {.notes}
**How to read this plot:**
- Dashed lines = NoPriority baseline (72.6s bus, 36.2s car)
- Green quadrant = BOTH bus AND car better than baseline (the Pareto zone)
- Point size = grant rate (bigger = more TSP actions)
- Color = total delay (darker = worse)

**Key story:**
- WaveGate, CellSearch, and NashGate are all in the green zone — both bus AND
  car improve. NashGate has the largest total-delay reduction ($-$18.5\%) and
  the lowest car delay (32.7s); CellSearch is close behind ($-$17.9\%) with
  the largest bus-delay reduction of the three ($-$31.7\%).
- CPD-QL and CellQLearn are both in the "both worse" zone (top-right) — neither
  guard has converged to a useful policy.
:::

---

# Z₁/Z₂ Weights: Why Z₂ Is Always Zero (+ New Z₄ Term)

```{r}
wobj <- df %>% filter(str_detect(experiment,"WOBJ_SWEEP"))
wobj %>% group_by(beta) %>% summarise(Total=mean(pass_delay_hrs), Z1=mean(wobj_Z1_total),
    Z2=mean(wobj_Z2_total), J=mean(wobj_objective_total), .groups="drop") %>%
  mutate(Delta=Total-base_delay, Label=paste0("β=",beta)) %>%
  bind_rows(tibble(Label="Baseline",Total=base_delay,Delta=0,Z1=NA,Z2=NA,J=NA)) %>%
  select(Label, Total, Delta, Z1, Z2, J) %>% arrange(Total) %>%
  gt() %>% tab_header(title="Effect of Progression Weight β on Z₁/Z₂/J",
    subtitle="J = α·Z₁ + β·Z₂, α=0.8 — Z₂ is logged as 0.0 for every row") %>%
  fmt_number(columns=c(Total,Delta,Z1,Z2,J),decimals=1) %>%
  cols_label(Label="", Total="Total (h)", Delta="Δ (h)", Z1="Z₁ (pax·s)", Z2="Z₂ (pax·s)", J="J (pax·s)") %>%
  tab_options(table.font.size="85%", data_row.padding="4px") %>%
  data_color(columns=Delta, fn=function(x) scales::col_numeric("RdYlGn",domain=range(x,na.rm=TRUE),reverse=TRUE)(x))
```

::: {.notes}
**What is Z₂, concretely?**
Z₂ is the flow-weighted progression-bandwidth mismatch (Eq. Z2 in the paper) —
it measures how badly bus arrivals miss the "green wave" along the corridor.
For each coordinated link (i,j): τ_ij is the free-flow travel time from i to
j (how long a bus takes to drive between the two intersections at uncongested
speed), and φ_i/φ_j are the *realised* offsets at which each intersection's
bus-progression phase turns green. Compare the realised gap (φ_j − φ_i) to
the ideal gap τ_ij:
- If the realised gap is **larger** than τ_ij (signal j turns green later
  than a bus leaving i on green would need), the bus arrives at j too EARLY
  and must wait — this shortfall is **ε⁺**.
- If the realised gap is **smaller** than τ_ij, signal j's green has already
  started/finished before the bus arrives — the bus is too LATE and misses
  its window — this is **ε⁻**.
Either way, the green time actually usable by the bus shrinks by (ε⁺+ε⁻).
Z₂ sums (ε⁺+ε⁻) across every coordinated link/timestep, weighted by bus flow
q — i.e. it's the total "green-wave miss" time, in pax·s. A perfectly
synchronized green wave (every signal's offset exactly equal to the
inter-section free-flow travel time) gives ε⁺=ε⁻=0 and Z₂=0 — the
theoretical best case.

**Why is Z₂ always 0.0 here — and NOT because the green wave is perfect?**
Reaching Z₂=0 the way above requires the controller to track the realised
offsets φ_i at runtime and feed the resulting ε⁺/ε⁻ into the Z₂ formula.
None of the six methods in this dataset populate this signal (the
corridor-offset error `_last_offset_error_s` is never computed) — a telemetry
gap, not a property of any particular β, and not evidence of perfect
progression. So `J = α·Z₁ + β·Z₂` collapses to `J = α·Z₁` exactly for
every row: e.g. β=0.1, α=0.8 → 0.8 × 182,170.5 ≈ 145,736.4 ✓ (matches J).
β therefore has NO effect on J while Z₂≡0.

**So why does β still change the results?**
β is ALSO propagated directly into two reward weights:
`REWARD_MAIN_SECTION_WEIGHT = max(0.1, 1−β)` and
`REWARD_SIDE_SECTION_WEIGHT = 3β`. These reweight the MAIN-corridor vs.
SIDE-street components of Z₁ itself — a second, independent channel from Z₂.
This is the real mechanism behind the β results below.

**A new Z₄ term has just been added (results pending).**
A new `WOBJ_DELTA` weight (δ) extends the objective to
`J = α·Z₁ + β·Z₂ + δ·Z₄`, where Z₄ is the *raw* (unweighted: ρ=1, w_main=w_side=1)
sum of vehicle-delay-seconds across the same four section/mode components as
Z₁ — i.e. the quantity that drives Δ Avg TT (Total Vehicle-Hours slide), now
expressed as a per-step cost the reward can penalise directly. δ=0 (the
default) recovers `J = α·Z₁ + β·Z₂` exactly, so every row on this slide is
unaffected. A 4-config pending sweep (δ ∈ {0,2,8,32} at `D20/P0/T60/S2.5`,
the new overall-best configuration from the lateness-takeover follow-up
sweep) is described a few slides ahead.

**Key story:**
- β=0.1: low side-street weight → solver ignores cross-traffic →
  +15% worse than baseline.
- β=0.2: −25% vs baseline. β=0.3: best result (−28%).
- β≥0.2 is REQUIRED for any benefit. Below that, TSP does more harm than good.
- **Z₁ note**: β=0.2 has HIGHER Z₁ (357k) than β=0.3 (172k) because the
  reweighted objective makes different argmin choices that increase raw
  delay (Z₁) while still reducing the *Total* pax·hrs metric.
:::

---

# Z₃: META-TSP Gate Parameter Sweep (BUS_PRIO, 12 configs)

All 12 rows below use `GLOBAL_REWARD + META_TSP_MODE` — the "META-TSP" method
(see Solution Methods) — sweeping its three gate parameters
(`δ_min, n_min, τ`). This *is* the BUS_PRIO sweep. A follow-up 12-config
sweep of `τ × META_HW_SENSITIVITY`, anchored at this sweep's best config
(`D20/P0/T60`), is presented two slides ahead — with real results.

```{r}
busprio_lat <- busprio_real %>% rowwise() %>% mutate(
    bus_trips_file = paste0("results/", experiment, "_seed300_11129236_11129237_11129240/bus_trips.csv"),
    Pct30 = {
      tt <- read_csv(bus_trips_file, show_col_types=FALSE)$TravelTime_s
      tt <- tt[!is.na(tt) & tt>0]
      round(100*mean((tt-sched_tt)>30),1)
    }
  ) %>% ungroup() %>%
  transmute(
    min_lat = case_when(str_detect(experiment,"D0_")~"0", str_detect(experiment,"D10_")~"10", str_detect(experiment,"D20_")~"20"),
    min_ben = case_when(str_detect(experiment,"P0_")~"0", str_detect(experiment,"P40_")~"40"),
    takeover = case_when(str_detect(experiment,"T30")~"30", str_detect(experiment,"T60")~"60"),
    Total = pass_delay_hrs, Bus = bus_delay_s, Car = car_delay_s,
    AvgBusTT = stats_AvgBusTT_s, MeanSigma = round(stats_AvgBusTT_s - sched_tt,1), Pct30,
    TotalTT = stats_Net_TotalTT_h_Bus+stats_Net_TotalTT_h_Car+stats_Net_TotalTT_h_Truck,
    TotalVeh = stats_N_DistinctCars+stats_N_DistinctBuses+stats_N_DistinctTrucks,
    DeltaAvgTT = round(TotalTT*3600/TotalVeh - base_avgtt, 1)
  ) %>% select(-TotalTT,-TotalVeh) %>% arrange(Total)

busprio_lat %>%
  bind_rows(tibble(min_lat="—", min_ben="—", takeover="—", Total=base_delay,
    Bus=base_bus, Car=base_car, AvgBusTT=no_tsp$stats_AvgBusTT_s[1],
    MeanSigma=round(no_tsp$stats_AvgBusTT_s[1]-sched_tt,1),
    Pct30=round(mean(lateness>30)*100,1), DeltaAvgTT=0)) %>%
  gt() %>% tab_header(title="META-TSP Gate Parameter Sweep",
    subtitle="δ_min = min bus delay, n_min = min net benefit, τ = lateness-takeover threshold") %>%
  fmt_number(columns=c(Total,Bus,Car,AvgBusTT,MeanSigma,Pct30,DeltaAvgTT), decimals=1) %>%
  cols_label(min_lat="δ_min (s)", min_ben="n_min (pax·s)", takeover="τ (s)", Total="Total (h)",
    Bus="Bus (s)", Car="Car (s)", AvgBusTT="Avg Bus TT (s)", MeanSigma="Mean σ (s)", Pct30="σ>30s %",
    DeltaAvgTT="Δ Avg TT (s)") %>%
  tab_options(table.font.size="70%", data_row.padding="2px") %>%
  data_color(columns=Total, fn=function(x) scales::col_numeric("RdYlGn",domain=range(x,na.rm=TRUE),reverse=TRUE)(x)) %>%
  data_color(columns=DeltaAvgTT, fn=function(x) scales::col_numeric("RdYlGn",domain=range(x,na.rm=TRUE),reverse=TRUE)(x)) %>%
  tab_source_note("D20/P0/T60 and D20/P40/T60 are numerically identical (n_min isn't binding once δ_min=20 already gates most buses); same for the D0/T30 vs T60 and D0/P40/T30 vs T60 pairs. Δ Avg TT (last row = NoPriority, by definition 0) is the change in unweighted average travel time per vehicle — shown for context, not part of the optimised objective.")
```

::: {.notes}
**What each parameter does:**
- **δ_min (Min Lateness)**: bus must already be ≥ this delayed before META-TSP
  considers acting. 0 = help any bus. 20 = only buses already meaningfully late.
- **n_min (Min Benefit)**: the CTM cost reduction `Z(NO_ACTION)-Z(a*)` must be
  ≥ this many net pax·s. 0 = any saving counts. 40 = only substantial savings.
- **τ (Takeover)**: if `|σ_in| ≥ τ`, both gates above are bypassed — META-TSP
  fires unconditionally.

**Key story:**
- **D20/P0/T60 is the standout result on passenger delay within this
  12-config sweep**: −24.5% total delay (952.4h vs NoPriority's 1261.1h),
  with both bus (−31.4%) and car (−19.0%) improving — a larger Pareto gain
  than even CellSearch. A follow-up sweep (two slides ahead), varying
  `META_LATENESS_TAKEOVER_S × META_HW_SENSITIVITY` at this same `D20/P0`
  anchor, finds an even better configuration — `D20/P0/T60/S2.5`, −27.2%,
  the new overall best of all 60 executed experiments.
- **But its Mean σ (36.9s) is still higher than NoPriority's (35.5s)** — like
  every configuration tested (including the follow-up sweep's new best), it
  does not beat the no-TSP baseline on per-passage schedule lateness, even
  though it is the best on total delay. Among configs that improve total
  delay by >100 pax·hrs, D20/P0/T60 has the smallest σ gap (+1.4s) other than
  CellSearch's (+0.7s).
- **D20/P0/T60 is also not the standout result on Δ Avg TT within this
  sweep.** Its +34.1s is *worse* than CellSearch (+21.9s), WaveGate
  (+25.5s), or NashGate (+28.9s) — despite its best-in-sweep passenger delay.
  Within this sweep, D10/P40/T60 has the smallest Δ Avg TT (+26.7s) but a
  much smaller delay improvement (−70.2 pax·hrs vs. D20/P0/T60's −308.7). The
  follow-up sweep's `D20/P0/T60/S2.5` does slightly better still (+33.0s vs
  +34.1s) — a small step in the right direction, not a fix. Same pattern as
  the Z₁/Z₂ slide: Z₁ (occupancy-weighted) and Δ Avg TT (unweighted) don't
  always agree on which config is "best".
- **D0 configs (δ_min=0) are worst on every metric** — META-TSP fires for
  buses that are barely late, disrupting cross-traffic (Car delay up to
  54.2s) without a lateness payoff (σ>30s up to 56.5%, worse than baseline).
- The follow-up sweep (two slides ahead) presents real results for the
  τ × sensitivity sweep aimed at closing the σ gap without making Δ Avg TT
  worse.
:::

---

# Bus Lateness by Strategy (Real Data)

```{r}
lat_dirs <- list(NoPriority="results/NO_TSP_seed300_11129236_11129237_11129240/bus_trips.csv",
  CPD_QL="results/DCTSP_MARL_seed300_11129236_11129237_11129240/bus_trips.csv",
  WaveGate="results/DCTSP_ZIG_seed300_11129236_11129237_11129240/bus_trips.csv",
  CellSearch="results/DCTSP_MP_ECTM_seed300_11129236_11129237_11129240/bus_trips.csv",
  CellQLearn="results/DCTSP_BXT_seed300_11129236_11129237_11129240/bus_trips.csv",
  NashGate="results/DCTSP_BARGAIN_SPM_seed300_11129236_11129237_11129240/bus_trips.csv",
  "META-TSP (best)"="results/LATENESS_SWEEP_T60_S25_seed300_11129236_11129237_11129240/bus_trips.csv")
sched <- as.numeric(quantile(read_csv(lat_dirs[["NoPriority"]],show_col_types=FALSE)$TravelTime_s,0.10))
lat_tbl <- tibble()
for(nm in names(lat_dirs)) { if(file.exists(lat_dirs[[nm]])) {
  tt <- read_csv(lat_dirs[[nm]],show_col_types=FALSE)$TravelTime_s; tt <- tt[!is.na(tt)&tt>0]
  lat_tbl <- bind_rows(lat_tbl, tibble(Strategy=ifelse(nm=="CPD_QL","CPD-QL",nm),
    Mean_s=round(mean(tt-sched),1), Pct30=round(100*mean(tt-sched>30),1),
    Pct60=round(100*mean(tt-sched>60),1)))
}}
lat_tbl %>% arrange(Mean_s) %>%
  gt() %>% tab_header(title="Actual Lateness Violation Rates", subtitle=paste0("Scheduled travel time: ",round(sched,1),"s (10th percentile free-flow)")) %>%
  cols_label(Strategy="", Mean_s="Mean σ (s)", Pct30=">30s %", Pct60=">60s %") %>%
  tab_options(table.font.size="90%", data_row.padding="4px") %>%
  data_color(columns=Mean_s, fn=function(x) scales::col_numeric("RdYlGn",domain=range(x,na.rm=TRUE),reverse=TRUE)(x))
```

::: {.notes}
**How to read:**
- Scheduled TT = 10th percentile of NoPriority travel times (free-flow estimate).
- σ = actual TT − scheduled, *per bus-intersection passage*. Positive = late.
- This is a different, smaller-scale metric from "Avg Bus Pass Delay" (the
  route-level delay-vs-desired-speed statistic in earlier slides) — TSP
  improves that metric substantially (e.g. 72.6→45.6s for CellSearch) while
  barely moving this one.

**Key story:**
- **NoPriority is lowest (35.5s / 43.3%>30s) of all seven rows** — every TSP
  strategy, including the best on total delay (CellSearch 36.2s, META-TSP
  `D20/P0/T60/S2.5` 37.4s, the new overall-best config), is slightly worse
  here.
- Why: when a TSP action shortens delay at intersection $i$, it can shift the
  downstream signal phase so a bus hits a red phase at $i+1$ it would
  otherwise have missed — a coordination cascade (see the BUS 11573 example:
  a 23.5s local saving costs 53s downstream). Averaged over many passages,
  these cascade losses slightly outweigh the local gains.
- CPD-QL is still the worst by far (53.8s / 59.2%>30s) — unguarded TSP
  disrupts the green wave on nearly every passage.
- **No strategy directly optimises σ.** The lateness-takeover ×
  adaptive-weight sweep (two slides ahead) is the first attempt aimed
  specifically at closing this gap — and, while it does not close it, it
  finds the new overall-best configuration on total passenger delay.
:::

---

# Z₃ Follow-up: Lateness-Takeover × Adaptive-Weight Sweep (12 configs, real data)

A second 12-config sweep, anchored at the previous slide's best gate region
(`δ_min=20, n_min=0`, i.e. `D20/P0`), varies `META_LATENESS_TAKEOVER_S` (τ)
∈ {60, 90, 120, 9999} × `META_HW_SENSITIVITY` (S) ∈ {0.5, 1.5, 2.5} — the
adaptive bus-weight slope that the BUS_PRIO sweep held fixed at its default
of 1.5:

$$w_{bus}(\sigma) = \texttt{META\_BASE\_BUS\_WEIGHT} + S \times
\mathrm{clip}\!\left(\frac{|\sigma_{in}|}{\texttt{META\_HW\_REF\_S}}, 0, 1\right)$$

τ = 9999 effectively disables the lateness-takeover gate.

```{r}
lat_sweep <- df %>% filter(str_detect(experiment, "^LATENESS_SWEEP")) %>%
  rowwise() %>% mutate(
    bus_trips_file = paste0("results/", experiment, "_seed300_11129236_11129237_11129240/bus_trips.csv"),
    tt_list = list({
      tt <- read_csv(bus_trips_file, show_col_types=FALSE)$TravelTime_s
      tt[!is.na(tt) & tt>0]
    })
  ) %>% ungroup() %>%
  mutate(
    tau  = case_when(str_detect(experiment,"_T60_")~"60", str_detect(experiment,"_T90_")~"90",
                     str_detect(experiment,"_T120_")~"120", str_detect(experiment,"_T9999_")~"9999"),
    sens = case_when(str_detect(experiment,"_S05$")~"0.5", str_detect(experiment,"_S15$")~"1.5", str_detect(experiment,"_S25$")~"2.5"),
    Total = pass_delay_hrs,
    DeltaTotal = round(100*(round(Total,1)-round(base_delay,1))/round(base_delay,1),1),
    Bus = bus_delay_s, Car = car_delay_s, AvgBusTT = stats_AvgBusTT_s,
    MeanSigma = map_dbl(tt_list, ~round(mean(.x - sched_tt),1)),
    Pct30 = map_dbl(tt_list, ~round(100*mean((.x-sched_tt)>30),1)),
    TotalTT = stats_Net_TotalTT_h_Bus+stats_Net_TotalTT_h_Car+stats_Net_TotalTT_h_Truck,
    TotalVeh = stats_N_DistinctCars+stats_N_DistinctBuses+stats_N_DistinctTrucks,
    DeltaAvgTT = round(TotalTT*3600/TotalVeh - base_avgtt, 1)
  ) %>%
  select(tau, sens, Total, DeltaTotal, Bus, Car, AvgBusTT, MeanSigma, Pct30, DeltaAvgTT) %>%
  arrange(Total)

lat_sweep %>%
  bind_rows(tibble(tau="—", sens="—", Total=base_delay, DeltaTotal=0,
    Bus=base_bus, Car=base_car, AvgBusTT=no_tsp$stats_AvgBusTT_s[1],
    MeanSigma=round(no_tsp$stats_AvgBusTT_s[1]-sched_tt,1),
    Pct30=round(mean(lateness>30)*100,1), DeltaAvgTT=0)) %>%
  gt() %>% tab_header(title="Lateness-Takeover × Adaptive-Weight Sweep (anchored at D20/P0)",
    subtitle="τ = META_LATENESS_TAKEOVER_S (s), S = META_HW_SENSITIVITY") %>%
  fmt_number(columns=c(Total,DeltaTotal,Bus,Car,AvgBusTT,MeanSigma,Pct30,DeltaAvgTT), decimals=1) %>%
  cols_label(tau="τ (s)", sens="S", Total="Total (h)", DeltaTotal="ΔTotal %",
    Bus="Bus (s)", Car="Car (s)", AvgBusTT="Avg Bus TT (s)", MeanSigma="Mean σ (s)",
    Pct30="σ>30s %", DeltaAvgTT="Δ Avg TT (s)") %>%
  tab_options(table.font.size="70%", data_row.padding="2px") %>%
  data_color(columns=Total, fn=function(x) scales::col_numeric("RdYlGn",domain=range(x,na.rm=TRUE),reverse=TRUE)(x)) %>%
  data_color(columns=DeltaAvgTT, fn=function(x) scales::col_numeric("RdYlGn",domain=range(x,na.rm=TRUE),reverse=TRUE)(x)) %>%
  tab_source_note("τ∈{90,120,9999} rows are numerically identical across all three sensitivity values — the lateness-takeover gate rarely fires once τ≥90, so the δ_min/n_min cost-benefit gate governs and is insensitive to META_HW_SENSITIVITY. τ=60/S=1.5 reproduces BUSPRIO_SWEEP_D20_P0_T60 exactly (cross-check). Last row = NoPriority.")
```

::: {.notes}
**New overall best: `D20/P0/T60/S2.5`.** Two of the twelve configurations —
τ=60 at S=0.5 and S=2.5 — improve on `D20/P0/T60`. `D20/P0/T60/S2.5`
(`LATENESS_SWEEP_T60_S25`) reaches 918.1 pax·hrs (−27.2% vs NoPriority), the
lowest total passenger delay of all 60 executed experiments, with
Pareto-improving bus (−33.9%) and car (−22.8%) delay — both *larger*
improvements than `D20/P0/T60`'s (−31.4%/−19.0%). `D20/P0/T60/S0.5` is a
close second (928.2 pax·hrs, −26.4%). For τ≥90, total delay gets
monotonically worse as τ increases (959.0 → 1004.7 → 1070.6 pax·hrs at
τ=90,120,9999) — relaxing the takeover threshold beyond 60s costs passenger
delay regardless of sensitivity.

**Mean σ goal not met — and the new best is no exception.** None of the 12
configurations push Mean σ below NoPriority's 35.5s; the closest is
`D20/P0/T60` itself (36.9s). `D20/P0/T60/S2.5` (37.4s) and `D20/P0/T60/S0.5`
(37.8s) are both *slightly worse* on σ than the configuration they beat on
total delay — the adaptive-weight tuning that reduces total passenger delay
further does so at a small further cost in per-passage schedule adherence.
Within the (τ, sensitivity) region tested here, this mechanism cannot push
Mean σ below the NoPriority baseline; Z₃ would need a different formulation
(e.g. a direct σ-penalty term, future work) to do so.

**Δ Avg TT: a partial improvement.** `D20/P0/T60/S2.5`'s Δ Avg TT (+33.0s)
is marginally better than `D20/P0/T60`'s (+34.1s), but both remain worse
than CellSearch (+21.9s), WaveGate (+25.5s), and NashGate (+28.9s). The τ≥90
configurations have the *best* Δ Avg TT in this sweep (+27.9s at τ=90,
+27.7s at τ=120) — close to D10/P40/T60's +26.7s (the best-in-BUS_PRIO
value, previous slide) — but cost 41–87 pax·hrs more total delay than
`D20/P0/T60/S2.5`. No configuration in either sweep combines the lowest
total delay with the lowest Δ Avg TT — this remains the motivation for the
raw travel-time sweep (next slide), now re-anchored at `D20/P0/T60/S2.5`.

**Sensitivity is inert except at τ=60.** For τ∈{90,120,9999}, all three
sensitivity values produce numerically identical results — the adaptive
bus-weight slope has no observable effect. A plausible explanation: the
lateness-takeover condition `|σ_in| ≥ τ` fires often enough at τ=60 to
interact with `w_bus(σ)` during action selection, but rarely or never for
τ≥90 at this corridor's demand level — so those configurations are governed
entirely by the δ_min=20/n_min=0 cost-benefit gate, whose argmin appears
insensitive to `w_bus` in the range tested.

**Built-in cross-check**: τ=60/S=1.5 reproduces `BUSPRIO_SWEEP_D20_P0_T60`
exactly (952.4 pax·hrs, Mean σ=36.9s, Δ Avg TT=+34.1s) — confirmed.
:::

---

# Next: Raw Travel-Time Term (Z₄)

**Results pending** — requires a follow-up Aimsun batch run.

A second new term, $Z_4$, is added directly to the objective: the *raw,
unweighted* sum of vehicle-delay-seconds across all four section/mode
components — the same quantity that drives Δ Avg TT (Total Vehicle-Hours
slide), now expressed as a per-step cost the reward can minimise directly.

$$J = \alpha Z_1 + \beta Z_2 + \delta Z_4, \qquad \delta = \texttt{WOBJ\_DELTA}$$

Anchored at the new overall-best configuration (`D20/P0/T60/S2.5`, i.e.
`LATENESS_SWEEP_T60_S25` — the −27.2% total-delay result from the previous
slide), a new 4-config sweep varies δ:

| `WOBJ_DELTA` (δ) | Rationale |
|---|---|
| 0 | Reproduces `D20/P0/T60/S2.5` exactly — built-in cross-check |
| 2, 8 | Comparable to Z₁'s car-delay weight (α·ρ_car ≈ 1.2) |
| 32 | Comparable to Z₁'s bus-delay weight (α·ρ_bus ≈ 32) |

**Goal**: `D20/P0/T60/S2.5` is the lowest-total-delay configuration of all 60
executed experiments but its Δ Avg TT (+33.0s) is still *worse* than
CellSearch's +21.9s. A non-zero δ should pull the solver away from this
trade-off, reducing Δ Avg TT towards CellSearch's level without giving up
`D20/P0/T60/S2.5`'s passenger-delay advantage over `D20/P0/T60` (−27.2% vs
−24.5%).

::: {.notes}
This 4-config sweep (`TRAVELTIME_SWEEP_D{0,2,8,32}`) is enabled in
`batch_runner.py`, re-anchored this round from `D20/P0/T60` to
`D20/P0/T60/S2.5` (`META_LATENESS_TAKEOVER_S=60, META_HW_SENSITIVITY=2.5`),
but has not been run yet — like the lateness-takeover × adaptive-weight
sweep on the previous slide, it requires Aimsun and is a separate batch job.
This is the only experiment of the 64 designed that hasn't been run yet.

The mechanism is implemented in `compute_weighted_objective` /
`_accumulate_weighted_objective` (`intersection_controller.py`): Z₄ is
computed and logged alongside Z₁/Z₂ in the same per-run weighted-objective
trace, with `WOBJ_DELTA` defaulting to 0.0 (legacy behaviour — every existing
result on the previous slides is unaffected).

The lateness-takeover × adaptive-weight sweep (previous slide) already
improved Δ Avg TT slightly (+34.1s → +33.0s) while improving total delay
substantially (−24.5% → −27.2%) — a step in the right direction on both
fronts, but not a fix for the Z₁ / Δ Avg TT gap. This sweep targets that gap
directly via δ, from the new, better-positioned anchor.
:::

---

# Per-Objective Sensitivity: Best Config vs. Baseline

The previous slides report each term of $J = \alpha Z_1 + \beta Z_2 + \gamma Z_3 + \delta Z_4$
separately, across different tables. Here they are collected for a single
configuration — **`D20/P0/T60/S2.5`** (`LATENESS_SWEEP_T60_S25`), the
configuration with the lowest total passenger delay of all 60 executed
experiments — against the **NoPriority** baseline.

| Objective | Metric | NoPriority | D20/P0/T60/S2.5 | Δ | Change |
|---|---|---|---|---|---|
| $Z_1$ (passenger delay) | PaxDelay (pax·hrs) | 1261.1 | 918.1 | −343.0 | **−27.2%** |
| $Z_2$ (offset correction) | \|error\| (pax·s) | 0.0 | 0.0 | 0.0 | n/a† |
| $Z_3$ (schedule lateness) | Mean σ (s) | 35.5 | 37.4 | +1.9 | +5.4% |
| $Z_4$ (raw travel-time delay) | Δ Avg TT (s/veh) | 0.0 (ref.) | +33.0 | +33.0 | ‡ |

† $Z_2 \equiv 0$ for every configuration in this dataset (telemetry gap, see Z₁/Z₂ slide).
‡ $Z_4$ itself has no logged data yet (pending sweep, previous slide); Δ Avg TT is the
best available proxy for its sign and magnitude.

::: {.notes}
`D20/P0/T60/S2.5` was selected purely on total passenger delay (previous two
slides), and on that axis it is decisive: −27.2% passenger delay, the best
of all 60 executed experiments. $Z_2$ is unaffected — still exactly zero, a
telemetry gap rather than a result.

On $Z_3$ it is *worse* than the baseline by 1.9s of mean schedule lateness
(+5.4%) — a slightly larger gap than `D20/P0/T60`'s (+1.4s, +3.9%) — though
still better than NashGate, CellQLearn, and CPD-QL, and better than
`D20/P0/T60/S0.5` and all τ≥90 configurations in the lateness-takeover ×
adaptive-weight sweep.

On the $Z_4$ proxy (Δ Avg TT) it is, again, the *worst* Pareto-improving
result in the entire study: +33.0s/veh — marginally better than
`D20/P0/T60`'s +34.1s, but still well above +21.9 for CellSearch, +25.5 for
WaveGate, and +28.9 for NashGate.

**Bottom line**: the configuration that wins on total passenger delay remains
close to a wash on $Z_3$ and a clear loss on $Z_4$ relative to the other
Pareto-improving strategies — the motivation for the raw travel-time sweep
(previous slide), now re-anchored at `D20/P0/T60/S2.5`.
:::

---

# Delay Split: Main Corridor vs Side Streets

```{r fig.height=4}
strategies %>% select(strategy_label,main_delay_hrs,side_delay_hrs) %>%
  pivot_longer(-strategy_label,names_to="section",values_to="hrs") %>%
  mutate(section=ifelse(section=="main_delay_hrs","Main Corridor","Side Streets")) %>%
  ggplot(aes(reorder(strategy_label,hrs),hrs,fill=section)) +
  geom_col(position="stack",width=0.5) + coord_flip() +
  scale_fill_manual(values=c("Main Corridor"="#8e44ad","Side Streets"="#2ecc71")) +
  labs(x="",y="Passenger Delay (hrs)",fill="")
```

::: {.notes}
**How to read:**
- Purple = delay on the main corridor (KG Road — where buses travel).
- Green = delay on side streets (cross-traffic).
- Side streets dominate (69% of NoPriority total) because cross-traffic volume is high.

**Key story:**
- CPD-QL inflates MAIN corridor delay massively (purple bar grows) because its TSP
  actions disrupt the green wave → all traffic on KG Road suffers.
- WaveGate and NashGate reduce side-street delay while keeping main near baseline.
- The balance guard protects cross-traffic by blocking TSP when car cost is too high.
:::

---

# Key Findings

| Finding | Detail |
|---------|--------|
| **CPD-QL harmful** | +40% pax delay, +82% car veh·hrs, lateness WORSE than baseline (59% vs 43%) |
| **NashGate best Pareto** | −18.5% pax delay, −28.1% bus, −9.6% car — largest Pareto improvement of the six methods |
| **CellSearch nearly matches** | −17.9% pax delay via pure CTM search, no learning or bargaining — smallest per-vehicle delay increase (+21.9s) |
| **WaveGate best-tuned** | gate threshold 1.0, penalty 3.0 → −32% pax delay, −41% bus, −23% car |
| **β ≥ 0.2 required** | β=0.1 → +15% worse than baseline (progression ignored) |
| **D20/P0/T60/S2.5 is the new overall best** | −27.2% pax delay (918.1h), Pareto bus −33.9% / car −22.8% — lowest total delay of all 60 executed experiments |
| **No config beats baseline σ** | Even the new best (D20/P0/T60/S2.5) has Mean σ=37.4s > NoPriority's 35.5s, and its ΔAvgTT (+33.0s) is worse than CellSearch's (+21.9s) |
| **Total time metric** | CPD-QL +255.7 veh·hrs, WaveGate +52.1, NashGate +62.8 |

::: {.notes}
**Presenter summary:**

1. CPD-QL without a guard is worse than doing nothing.
2. NashGate and CellSearch are the two standout strategies — both improve bus
   AND car delay simultaneously (Pareto improvements), NashGate slightly
   ahead (−18.5% vs −17.9%).
3. CellSearch reaches this via pure CTM cost-minimisation with no learning or
   bargaining — remarkable given its simplicity. CellQLearn (same CTM model
   plus a Q-learning policy) is the second-worst strategy, showing the
   learned policy hasn't converged to anything useful.
4. Progression weight β must be ≥0.2 or TSP makes things worse — and its
   effect operates via the reward-section weights, not via Z₂ (Z₂≡0 here).
5. The lateness-takeover × adaptive-weight sweep found `D20/P0/T60/S2.5`
   (−27.2% total delay), the new overall best of all 60 executed
   experiments — ahead of `D20/P0/T60` (−24.5%, the best of the 12-config
   BUS_PRIO sweep) and every strategy default.
6. Schedule-lateness (σ) remains unsolved: even `D20/P0/T60/S2.5`
   (Mean σ=37.4s) doesn't beat the no-TSP baseline (35.5s) on this metric. A
   direct σ-penalty formulation is flagged as future work.
7. Total vehicle-hours is the simplest metric: CPD-QL +255.7 veh·hrs,
   WaveGate +52.1, NashGate +62.8 (now includes bus+car+truck hours).
:::

---

# Experiment Plan

**64 experiments designed, 60 executed, single seed 300:**

| Group | Count | What it tests |
|-------|-------|---------------|
| Strategy comparison | 6 | NoPriority, CPD-QL, WaveGate, CellSearch, CellQLearn, NashGate |
| WOBJ sweep | 9 | α (delay weight) × β (progression weight) — Z₁ vs Z₂ |
| Main-Side sweep | 9 | w_main × w_side — corridor vs cross-street priority |
| WaveGate variants | 12 | Strategy configuration variants |
| **META-TSP gate sweep (BUS_PRIO)** | **12** | **δ_min × n_min × τ — real data** |
| **Lateness-takeover × adaptive-weight sweep** | **12** | **τ × META_HW_SENSITIVITY @ D20/P0 — real data, new overall best** |
| **Raw travel-time sweep** | **4** | **δ (WOBJ_DELTA) ∈ {0,2,8,32} @ D20/P0/T60/S2.5 — *pending*** |
| **Total** | **64 (60 executed)** | |

::: {.notes}
**What each sweep tells us:**
- WOBJ: how much should we weight progression (Z₂) vs delay (Z₁)? In
  practice β's effect comes via the reward-section weights, not Z₂ — see the
  Z₁/Z₂ Weights slide.
- Main-Side: should we protect the corridor or cross-streets more?
- META-TSP gate sweep (BUS_PRIO): when should the lateness-aware gate fire
  for late buses regardless of car cost? Produced the lowest passenger delay
  of the 12-config BUS_PRIO sweep (D20/P0/T60, −24.5% total delay) — though
  not the lowest Δ Avg TT (CellSearch, WaveGate, and NashGate all beat it on
  that axis).
- Lateness-takeover × adaptive-weight sweep (12 configs, real data):
  anchored at D20/P0, varies τ ∈ {60,90,120,9999} × META_HW_SENSITIVITY ∈
  {0.5,1.5,2.5}. Found `D20/P0/T60/S2.5` (−27.2%), the new overall best of
  all 60 executed experiments — see the dedicated slide. Targets the
  unsolved σ-lateness gap from the previous slides while tracking Δ Avg TT
  as a secondary goal.
- Raw travel-time sweep (4 configs, pending): adds the new Z₄ term
  (Section "Next: Raw Travel-Time Term") to the reward via
  `WOBJ_DELTA` ∈ {0,2,8,32}, re-anchored this round at
  `D20/P0/T60/S2.5` (`LATENESS_SWEEP_T60_S25`). Targets its Δ Avg TT gap
  (+33.0s vs. CellSearch's +21.9s) directly. Requires a follow-up Aimsun
  batch run (results pending) — the only experiment of the 64 designed that
  hasn't been run yet.
:::

# Thank You
